Supplemental Materials for ``Universals of word order result from optimization of grammars for efficient communication''

### 2019-05-18 second-draft/pnas-word-order-si.tex
SI

In this section, we describe the method we employ for creating grammars that are optimized for efficiency.
We carry out optimization in an extended space of grammars that interpolates continuously between different grammars (Section~\ref{sec:diff-gramm}).
More specifically, we include probabilistic relaxations of grammars, which describe probability distributions over different ways of ordering a syntactic structure into a sentence.
This makes efficiency a \emph{differentiable} function of the grammar parameters, and enables efficient optimization with stochastic gradient descent, as we describe in Section~\ref{sec:optim-eff}.

@@

At the beginning of the optimization procedure, we initialize all values $a_\tau := 0.5$, $b_\tau := 0$ (except for the \emph{obj} dependency, for which we fix $a_\tau$ to 0 or 1, see Section~\ref{sec:neural-architectures}).

See Section~\ref{sec:neural-architectures} for the stopping criterion and learning rates.

>>>>

In this section, we describe the method we employ for creating grammars that are optimized for efficiency.
%Our method will be applicable more generally to the task of optimizing grammars for a broad class of objective functions.
We carry out optimization in an extended space of grammars that interpolates continuously between different grammars (Section~\ref{sec:diff-gramm}).
More specifically, we include probabilistic relaxations of grammars, which describe probability distributions over different ways of ordering a syntactic structure into a sentence.
This makes efficiency a \emph{differentiable} function of the grammar parameters, and enables efficient optimization with stochastic gradient descent, as we describe in Section~\ref{sec:optim-eff}.

@@

At the beginning of the optimization procedure, we initialize all values $a_\tau := 0.5$, $b_\tau := 0$ (except for the \emph{obj} dependency, for which we fix $a_\tau$ to 0 or 1, see Section~\ref{sec:neural-architectures}).
Empirically, we observe that optimizing differentiable ordering grammars for efficiency leads to convergence towards deterministic behavior, allowing us to extract equivalent deterministic grammars as described in Section~\ref{sec:diff-gramm}.

See Section~\ref{sec:neural-architectures} for the stopping criterion and learning rates used in this optimization scheme.
